---
title: "Tanzania Economic Indicators — Automated Report"
author: "Generated by the ETL pipeline"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
  # Uncomment to render PDF instead (requires a LaTeX installation, e.g. tinytex):
  # pdf_document:
  #   toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 9)

library(DBI)
library(RSQLite)
library(dplyr)
library(tidyr)
library(ggplot2)

DB_PATH <- file.path("..", "data", "tanzania_indicators.db")
if (!file.exists(DB_PATH)) DB_PATH <- file.path("data", "tanzania_indicators.db")

con <- dbConnect(RSQLite::SQLite(), DB_PATH)
df <- dbReadTable(con, "indicators")
dbDisconnect(con)
```

## Overview

This report is generated automatically from the SQLite database produced by
the Python ETL pipeline (`src/pipeline.py`), which pulls data from the
[World Bank Open Data API](https://data.worldbank.org/). It covers
`r nrow(df)` observations across `r length(unique(df$indicator))` indicators
for Tanzania, spanning `r min(df$year)`–`r max(df$year)`.

Re-running the pipeline and re-knitting this document (`Rscript -e
"rmarkdown::render('r_analysis/report.Rmd')"`) refreshes every number and
chart below with no manual editing required — that's the point of the
automation.

## GDP growth over time

```{r gdp-growth}
gdp <- df %>% filter(indicator == "gdp_growth_pct")

ggplot(gdp, aes(x = year, y = value)) +
  geom_col(fill = "#2c7fb8") +
  geom_hline(yintercept = 0, linewidth = 0.3) +
  labs(
    title = "Tanzania GDP Growth (annual %)",
    x = NULL, y = "GDP growth (%)"
  ) +
  theme_minimal(base_size = 12)
```

## Inflation vs. GDP growth

```{r inflation-vs-gdp}
compare_df <- df %>%
  filter(indicator %in% c("gdp_growth_pct", "inflation_pct")) %>%
  select(year, indicator, value)

ggplot(compare_df, aes(x = year, y = value, color = indicator)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(
    title = "GDP Growth vs. Inflation",
    x = NULL, y = "Percent (%)", color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```

## Population and internet adoption

```{r population-internet}
pop_net <- df %>%
  filter(indicator %in% c("population_growth_pct", "internet_users_pct")) %>%
  select(year, indicator, value)

ggplot(pop_net, aes(x = year, y = value, color = indicator)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Population Growth vs. Internet Adoption",
    x = NULL, y = "Percent (%)", color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```

## Summary statistics table

```{r summary-table}
df %>%
  group_by(indicator) %>%
  summarise(
    `Years covered` = paste(min(year), "-", max(year)),
    `Mean` = round(mean(value, na.rm = TRUE), 2),
    `SD` = round(sd(value, na.rm = TRUE), 2),
    `Latest value` = value[which.max(year)],
    .groups = "drop"
  ) %>%
  knitr::kable()
```

## Notes on methodology

- Data source: World Bank Open Data API, country code `TZA`.
- Missing years in the raw API response are dropped rather than imputed,
  to avoid presenting fabricated values as observed data.
- Year-over-year percentage change is computed per indicator during the
  Python transform stage and stored alongside the raw values.

---
*This report and the underlying pipeline are fully reproducible — see the
project README for setup instructions.*
